Q1: Are lake and terrestrial primary productivity coherent?

### 2019-05-21 AnalysisCode/ms1_analyses.Rmd
ms1 analyses v2 slightly better

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)

@@

```{r summarize coherences, echo=TRUE, cache=FALSE}
#short timescales
hist(coh.chlaXaccndvi$accndvicoh.ts1)
hist(coh.chlaXmaxndvi$maxndvicoh.ts1)

quantile(coh.chlaXaccndvi$accndvip.ts1)
quantile(coh.chlaXmaxndvi$maxndvip.ts1)

@@

#long timescales
hist(coh.chlaXaccndvi$accndvicoh.ts2)
hist(coh.chlaXmaxndvi$maxndvicoh.ts2)

quantile(coh.chlaXaccndvi$accndvip.ts2)
quantile(coh.chlaXmaxndvi$maxndvip.ts2)

@@

#all timescales
hist(coh.chlaXaccndvi$accndvicoh.ts3)
hist(coh.chlaXmaxndvi$maxndvicoh.ts3)

quantile(coh.chlaXaccndvi$accndvip.ts3)
quantile(coh.chlaXmaxndvi$maxndvip.ts3)

alpha=0.05
sum(coh.chlaXaccndvi$accndvip.ts3<alpha)/nrow(coh.chlaXaccndvi)
sum(coh.chlaXmaxndvi$maxndvip.ts3<alpha)/nrow(coh.chlaXmaxndvi)

print(coh.chlaXaccndvi$accndviphi.ts3[coh.chlaXaccndvi$accndvip.ts3<alpha]/pi)
print(coh.chlaXmaxndvi$maxndviphi.ts3[coh.chlaXmaxndvi$maxndvip.ts3<alpha]/pi)

@@

modvars.accndvi<-left_join(predictors, coh.chlaXaccndvi, by="lagoslakeid")
modvars.accndvi$nhd_ftype<-factor(modvars.accndvi$nhd_ftype)
modvars.accndvi$tsi.cat<-factor(modvars.accndvi$tsi.cat)
modvars.accndvi$tslength<-modvars.accndvi$end-modvars.accndvi$start + 1
```

```{r linear modelling, echo=TRUE, cache=FALSE}

@@

#short timescales
fm.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.coh.accndvi.st)
dredge.coh.accndvi.st<-dredge(fm.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

fm.p.accndvi.st<-gls(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.p.accndvi.st)
dredge.p.accndvi.st<-dredge(fm.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

fm.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.phi.accndvi.st)
dredge.phi.accndvi.st<-dredge(fm.phi.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
fm.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.coh.accndvi.lt)
dredge.coh.accndvi.lt<-dredge(fm.coh.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

fm.p.accndvi.lt<-gls(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.p.accndvi.lt)
dredge.p.accndvi.lt<-dredge(fm.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

fm.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(fm.phi.accndvi.lt)
dredge.phi.accndvi.lt<-dredge(fm.phi.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```

>>>>

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)

@@

```{r summarize coherences, echo=TRUE, cache=FALSE}
#short timescales
hist(coh.chlaXaccndvi$accndvicoh.ts1, main="Accumulated NDVI, short timescales", xlab="Coherence", ylab="Frequency")
hist(coh.chlaXmaxndvi$maxndvicoh.ts1, main="Maximum NDVI, short timescales", xlab="Coherence", ylab="Frequency")

quantile(coh.chlaXaccndvi$accndvicoh.ts1)
quantile(coh.chlaXmaxndvi$maxndvicoh.ts1)

@@

phicls<-c(-1,-.75,-0.25,0.25,0.75,1)

hist(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts1<0.2]/pi, main="Accumulated NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvip.ts1<0.2]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

#long timescales
hist(coh.chlaXaccndvi$accndvicoh.ts2, main="Accumulated NDVI, long timescales", xlab="Coherence", ylab="Frequency")
hist(coh.chlaXmaxndvi$maxndvicoh.ts2, main="Maximum NDVI, long timescales", xlab="Coherence", ylab="Frequency")

quantile(coh.chlaXaccndvi$accndvicoh.ts2)
quantile(coh.chlaXmaxndvi$maxndvicoh.ts2)

@@

hist(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts2<0.2]/pi, main="Accumulated NDVI, long timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvicoh.ts2>0.6]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

@@

modvars.accndvi<-left_join(predictors, coh.chlaXaccndvi, by="lagoslakeid")
modvars.accndvi$nhd_ftype<-factor(modvars.accndvi$nhd_ftype)
modvars.accndvi$tsi.cat<-factor(modvars.accndvi$tsi.cat)
modvars.accndvi$tslength<-modvars.accndvi$end-modvars.accndvi$start + 1

@@

modvars.accndvi.phist<-modvars.accndvi[modvars.accndvi$accndvip.ts1<0.2,]
modvars.accndvi.philt<-modvars.accndvi[modvars.accndvi$accndvip.ts2<0.2,]
```

```{r gls modelling, echo=TRUE, cache=FALSE}
#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

gls.p.accndvi.st<-gls(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.st)
dredge.p.accndvi.st<-dredge(gls.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

gls.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.phist) #remove ftype b/c only lakes
summary(gls.phi.accndvi.st)
dredge.phi.accndvi.st<-dredge(gls.phi.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

gls.p.accndvi.lt<-gls(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.lt)
dredge.p.accndvi.lt<-dredge(gls.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

gls.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.philt,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.phi.accndvi.lt)
dredge.phi.accndvi.lt<-dredge(gls.phi.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```

GLS models sucked, but inspection of results suggested that mixed effects models with random effects of region on the intercept may likely be an improvement.

```{r mixed modelling, echo=TRUE, cache=FALSE}
lme.coh.accndvi.st<-lmer(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.st)
dredge.coh.accndvi.st<-dredge(lme.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

lme.p.accndvi.st<-lmer(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.st)
dredge.p.accndvi.st<-dredge(lme.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

lme.phi.accndvi.st<-lmer(accndviphi.ts1 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.phist, na.action="na.fail")
summary(lme.phi.accndvi.st)
dredge.phi.accndvi.st<-dredge(lme.phi.accndvi.st, beta="sd") #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.st))

lme.coh.accndvi.lt<-lmer(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.lt)
dredge.coh.accndvi.lt<-dredge(lme.coh.accndvi.lt, beta="sd") #lake type is marginal improvement over intercept only (dAIC = 0.1)
print(head(dredge.coh.accndvi.lt))

lme.p.accndvi.lt<-lmer(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.lt)
dredge.p.accndvi.lt<-dredge(lme.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

lme.phi.accndvi.lt<-lmer(accndviphi.ts2 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.philt, na.action="na.fail")
summary(lme.phi.accndvi.lt)
dredge.phi.accndvi.lt<-dredge(lme.phi.accndvi.lt, beta="sd") #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.lt))

Adding random effects of huc2 region on intercept helped somewhat. Some possibilities are to add more detailed lake type classes, and to model 2-3 regions separately.
